Are we living in a depressive society?

### 2024-04-14 outputs/paper/paper.qmd
Update paper.qmd

1. **Hispanic and African American Subgroups:** Both subgroups exhibit relatively higer rates od depressive symptoms throughout the observed period compared to other groups. This may suggest a greater impact of socio-economic and possibly systemic factors affecting these communities disproportionately. The maximum rates observed for Hispanis and African Americans were 41.10% and 41.43%, respectively, with Hispanics showing a minimum rate significantly lower at 18.77%, suggesting more substantial fluctuatinos.

2. **Caucasian Subgroup:** This group generally shows lower symptom rates than the Hispanic and African American subgroups but follows a similar trend of fluctuations.

3. **Asian Subgroup:** Typically, this line shows the lowest symptom rates among the subgroups.

4. **Other (Non-Hispanic, Other Races and Multiple Races):** This category shows a unique trend with variability.

@@

Detailed Analysis of Trends by Education Level
@fig-educandtrend-graph presents the depressive disorder rates for several educational subgroups, including those with a college degree or higher, some college or an associate's degree, a high school diploma or GED, and less than a high school diploma. Each line represents the trend within a specific educational subgroup, offering a visual representation of how mental health issues vary based on educational attainment.

1. **College Graduate or Higher:** This subgroup typically shows the lowest rates of depressive symptoms, which might indicate better access to resources, stronger economic stability, or higher awareness and utilization of mental health services.

2. **Some College/Associate's Degree:** Individuals in this group exhibit slightly higher rates than those with higher education but still lower than the less educated groups. This might reflect a mid-level access to resources and varying levels of stress related to educational and career expectations.

3. **High School Diploma or GED:** This subgroup shows higher rates of depressive symptoms compared to those with more education. The increased rates could be linked to economic challenges, job market limitations, or the stress associated with lower social mobility.

4. **Less Than a High School Diploma:** This group consistently shows the highest rates of depressive symptoms, indicating significant vulnerabilities. Factors such as lower economic stability, limited access to healthcare, and higher life stressors likely contribute to these higher rates.

@@

What is done in this paper? 
What is something that we learn about the world? 
What is another thing that we learn about the world?

@@

African American and Hispanic group, Peaks within these subgroups from @fig-raceandtrend-graph might correspond to specific social or economic stressors that disproportionately affect these communities.
Caucasian group, The graph points to underlying factors that may influence mental health across different community settings.

Trends of the other group might reflect the diverse socio-economic and cultural backgrounds within this subgroup.

Discussion of Results of the education subgroup
The trends across educational subgroups underscore the strong correlation between educational attainment and mental health outcomes. Higher education often provides better opportunities for employment and income, which are crucial determinants of mental health. Conversely, lower educational attainment is closely associated with various life stressors that exacerbate mental health issues.

>>>>

1. **Hispanic and African American Subgroups:** Both subgroups exhibit relatively higer rates of depressive symptoms throughout the observed period compared to other groups. This may suggest a greater impact of socio-economic and possibly systemic factors affecting these communities disproportionately. The maximum rates observed for Hispanics and African Americans were 41.10% and 41.43%, respectively, with Hispanics showing a minimum rate significantly lower at 18.77%, suggesting more substantial fluctuatinos.

2. **Caucasian Subgroup:** While the Caucasian group shows lower symptom rates than the Hispanic and African American groups, it follows a similar pattern of fluctuations over time. The maximum rate for Caucasians was 35.17%, with an average rate of 27.92%, indicating that while the intensity may vary, the underlying factors influencing depressive symptoms are somewhat consistent across these groups.

3. **Asian Subgroup:** This group consistently shows the lowest symptom rates among the compared groups, with a maximum rate of 35.27% and a remarkably low minimum of 10.67%. The average rate stands at 22.74%, reflecting cultural, social, or possibly systemic differences in how mental health issues are reported, treated, or perceived within this community.

4. **Other (Non-Hispanic, Other Races and Multiple Races):** This category presents a unique trend line with variable rates, which could be indicative of the heterogeneous nature of this group. The highest recorded was 46.07%, and the lowest was 24.23%, with an average of 37.89%, suggesting significant variability possibly stemming from the diverse socio-economic and cultural backgrounds represented within this subgroup.

@@

@fig-educandtrend-graph presents the depressive disorder rates for several educational subgroups, including those with a college degree or higher, some college or an associate's degree, a high school diploma or GED, and less than a high school diploma. Each line represents the trend within a specific educational subgroup, offering a visual representation of how mental health issues vary based on educational attainment.

1. **Less Than a High School Diploma:** Individuals with less than an high school diploma experienced the highest depressive disorder rates throughout the period, with peaks reaching nearly 47%.

2. **Some College/Associate's Degree:** Individuals in this group exhibit slightly higher rates than those with higher education but still lower than the less educated groups. This group also showed a maximum rate close to 40%, averaging 26 - 27% rates.

3. **High School Diploma or GED:** For individuals holding a high school diploma or GED, the depressive rates were slightly lower but still significant, peaking at nearly 39%. It is showing similar pattern with **Some College/Associate's Degree** subgroups.

4. **College Graduate or Higher:** This subgroup typically shows the lowest rates of depressive symptoms. The Bachelor's degree or higher subgroup exhibited the lowest rates across all metrics, with a peak rate of just over 31% generally averaging 15% during the period.

@@

@@

Overall, the trends in depressive disorder rates by race and ethnicity @fig-raceandtrend-graph  underscore disparities in mental health experiences among different racial and ethnic groups in the U.S. Trends of the other group might reflect the diverse socio-economic and cultural backgrounds within this subgroup. These differences likely stem from a complex interplay of socio-economic factors, access to mental health services, cultural perceptions of mental health, and possibly systemic inequalities. The quantitative findings of maximum, minimum, and average rates for each subgroup add depth to our understanding of these disparities, highlighting the specific challenges and variabilities within each community.

Discussion of Results of the education subgroup

Factors such as lower economic stability, limited access to healthcare, and higher life stressors likely contribute to these higher rates. This group's high average rate suggests a consistent burden of depressive symptoms, potentially exacerbated by factors such as economic insecurity and limited access to mental health resources.

Those with some college or an Associate's degree also showed high variability in depressive disorder rates indicating that even some post-secondary education does not fully mitigate the risk of higher depressive symptoms. This might reflect a mid-level access to resources and varying levels of stress related to educational and career expectations.

High school diploma or GED, The increased rates could be linked to economic challenges, job market limitations, or the stress associated with lower social mobility. This reflects moderate but considerable mental health challenges within this educational subgroup.

Bachelor's degree or higher subgroup which might indicate better access to resources, stronger economic stability, or higher awareness and utilization of mental health services.

The trends across educational subgroups underscore the strong correlation between educational attainment and mental health outcomes. Higher education often provides better opportunities for employment and income, which are crucial determinants of mental health. Conversely, lower educational attainment is closely associated with various life stressors that exacerbate mental health issues.
